---
title: "Kalman ImputeTS"
output: html_notebook
---

Simplistic notebook to impute user time-series data using the imputeTS toolbox.

Imports:
```{r}
library(imputeTS) # Imputation package
library(reticulate) # Python interface package
library(stringr) # String modifications
library(forecast) # Forecasting time-series library
```
Get required files
```{r}
getwd() # Current working directory
data_path <- 'x_data/'
files <- list.files(path=data_path) # Necessary files that need to be imputed
```

Information about missing values
```{r}
source_python('utils.py')
sample_data <- read_pkl(file.path(data_path,files[1]))
```


```{r}
printmissingstats <- function(data){
# Print the statistics of missing values for each column in the test array
  for(i in 1:ncol(data)){
    stats <- statsNA(data[,i],bins=4,printOnly=FALSE)
  
    if(stats[2]>0){ # Only print if there are actually NaNs in the column
      cat('Column: ',i)
      statsNA(data[,i],bins=4,printOnly=TRUE)
    }
  }  
}

shape <- function(data){
  ncols <- ncol(data)
  nrows <- nrow(data)
  
  print(c(nrows,ncols))
}

printmissingstats(sample_data)
```
```{r}
for(i in 1:ncol(sample_data)){
  plotNA.distribution(sample_data[,i],colPoints='steelblue',colBackgroundMV='indianred2',main='Distribution of NAs',xlab='Time',ylab='Value',col='aliceblue')
}
```

Imputation test: na.kalman
```{r}
sample_data_kalman <- na.kalman(sample_data,model='StructTS',smooth=TRUE,type='trend')

# And checking effect of imputation
printmissingstats(sample_data_kalman) # Shows that the imputation works on all columns except those where there are no values at all (empty column)
```

Imputation effect on test
```{r}
for(i in 1:ncol(sample_data)){
  stats <- statsNA(sample_data[,i],bins=4,printOnly=FALSE)
  
  if(stats[2]>0){ # Only print if there are actually NaNs in the column
    plotNA.imputations(sample_data[,i],sample_data_kalman[,i],legend=TRUE,main='Visualization Imputed Values', xlab='Time',ylab='Value',colLines='aliceblue',colWithImputations='red',colWithNA='blue')
  }
}

```

Perform Kalman imputations and save new data
```{r}
folder <- 'ximpute_kalman'
for(file in files){
  fname <- unlist(strsplit(file,".",fixed=TRUE))[1][1] 
  data <- read_pkl(file.path(data_path,file))
  data_impute <- na.kalman(data,model='StructTS',smooth=TRUE,type='trend')
  file_path <- file.path(folder,paste(fname,'csv',sep='.')) # Without .pkl ending
  
  write.csv(data_impute,file=file_path)
  
}
```

Perform Seasonally-Decomposed (seadec) imputations and save new data
```{r}
# Remove seasonal component of the time series and imputes the de-seasonalized series and then re-adds them
# Going through the different algorithms: interpolation, mean, random, kalman, ma
folder <- 'ximpute_seadec_ma' 
for(file in files){
  fname <- unlist(strsplit(file,'.',fixed=TRUE))[1][1] 
  data <- read_pkl(file.path(data_path,file))
  data_impute <- na.seadec(data,algorithm='ma')
  file_path <- file.path(folder,paste(fname,'csv',sep='.')) # Without .pkl ending
  
  write.csv(data_impute,file=file_path)
  
}
```

Perform Seasonally-Split (seasplit) imputations and save new data
```{r}
# Split a dataset into seasons and then performs imputations
# Going through the different algorithms: interpolation, mean, random, kalman
folder <- 'ximpute_seasplit_kalman' 
for(file in files){
  fname <- unlist(strsplit(file,'.',fixed=TRUE))[1][1] 
  data <- read_pkl(file.path(data_path,file))
  data_impute <- na.seasplit(data,algorithm='kalman')
  file_path <- file.path(folder,paste(fname,'csv',sep='.')) # Without .pkl ending
  
  write.csv(data_impute,file=file_path)
  
}
```

Auto-Arima test
```{r}
sample_data_arima <- sample_data
for(i in 1:ncol(sample_data_arima)){
  data <- sample_data_arima[,i] # Get the required column
  na_count <- sum(is.na(data)) # Count NaNs in a column
  
  if(length(data)>na_count){ # Only impute if the column doesn't consist of all NaNs
    fit <- auto.arima(data,D=1) # Fit an ARIMA model to each feature in the column
    model <- fit$model
  
    # Kalman smooth
    kalman_run <- KalmanRun(data,model) # Pass data to the Kalman run feature
    
    id.na <- which(is.na(data)) # Which is a NaN  
    
    for(nan in id.na){
      sample_data_arima[nan,i] <- model$Z %*% kalman_run$states[nan,]
      
    }
    
  }
}
```

```{r}
for(i in 1:ncol(sample_data)){
  plotNA.imputations(sample_data[,i],sample_data_arima[,i],legend=TRUE,main='Visualization Imputed Values', xlab='Time',ylab='Value',colLines='aliceblue',colWithImputations='red',colWithNA='blue')
}
```

Perform ARIMA fits to obtain a state-space model and then Kalman impute and save new data
```{r}
folder <- 'ximpute_arima_kalman'
for(file in files){
  fname <- unlist(strsplit(file,".",fixed=TRUE))[1][1]
  data <- read_pkl(file.path(data_path,file))
  data_impute <- data  # Copy data
  
  for(i in 1:ncol(data_impute)){
    col_data <- data_impute[,i] # Required column
    na_count <- sum(is.na(col_data)) # Count NaNs in a column
    
    if(length(col_data)>na_count){  # Only impute if the column isn't just a NaN-column
      fit <- auto.arima(col_data,D=1) # Fit an ARIMA model to feature (each column) and force a seasonality
      model <- fit$model
      
      # Kalman Run
      kalman_run <- KalmanRun(col_data,model) # Pass data, model to the Kalman Run filter
      
      id.na <- which(is.na(col_data)) # Which is a NaN
      
      for(nan in id.na){
        data_impute[nan,i] <- model$Z %*% kalman_run$states[nan,]
      }
    }
  }
  file_path <- file.path(folder,paste(fname,'csv',sep='.')) # Without .pkl ending
  
  write.csv(data_impute,file=file_path)
  
}
```

Auto-Arima test
```{r}
sample_data_arima <- sample_data
for(i in 1:ncol(sample_data_arima)){
  data <- sample_data_arima[,i] # Get the required column
  na_count <- sum(is.na(data)) # Count NaNs in a column
  
  if(length(data)>na_count){ # Only impute if the column doesn't consist of all NaNs
    fit <- auto.arima(data,D=1) # Fit an ARIMA model to each feature in the column
    model <- fit$model
  
    # Kalman smooth
    kalman_run <- KalmanRun(data,model) # Pass data to the Kalman run feature
    
    id.na <- which(is.na(data)) # Which is a NaN  
    
    for(nan in id.na){
      sample_data_arima[nan,i] <- model$Z %*% kalman_run$states[nan,]
      
    }
    
  }
}
```